### 2016-05-04 writing/latex/wilf.tex
start edits r2 to wilf

Another question often asked in the field of permutation patterns is that
of Wilf-equivalence. Two patterns \(\pi\) and \(\sigma\) are said to be
Wilf-equivalent if their avoidance sets have the same size at each
length. More formally

\begin{definition}[Wilf-equivalence]
    Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}
    if for all \(k_{} \ge 0\), \(\size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\).

    Two sets of permutation patterns \(R\) and \(S\) are are
    \emph{Wilf-equivalent} if for all \(k_{} \ge 0\),
    \(\size{\av[k]{R}} = \size{\av[k]{S}}\).
\end{definition}

Wilf-equivalence is of interest as if two permutation classes are enumerated
in the same way then there should exist a bijection between them, and therefore
any other combinatorial object that they represent.

@@

There are a number of symmetries we can use when examining Wilf-equivalences
to reduce the amount of work, it can be easily seen that the reverse, complement
and inverse operations preserve enumeration, and therefore these classes are trivially
Wilf-equivalent.

\begin{align*}
    \rev{\textpattern{}{2,3,1}{}} &= \textpattern{}{1,3,2}{}\\
    \com{\textpattern{}{2,3,1}{}} &= \textpattern{}{2,1,3}{}\\
    \inv{\textpattern{}{2,3,1}{}} &= \textpattern{}{3,1,2}{}
\end{align*}

The group of symmetries on permutations is isomorphic to the dihedral group
of order \(8\), the group of symmetries of a square. \emph{Reverse-inverse} and
\emph{reverse} correspond tothe generators of the dihedral group.

Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
must only use these symmetries when they preserve the dominating pattern.

Throughout this section we will consider Wilf-equivalences of patterns
whilst avoiding the \emph{Dominating Pattern} \(\perm{2,3,1}\). We will
use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
be the usual Catalan generating function satisfying \(C(x) = 1 + C(x)^2\).
This is easy to see by structural decomposition around the maximum, as
shown in \FigureRef{fig:decompmax}

\begin{figure}[!ht]
    \centering
    \decompmax{0/1,1/0}{A}{}{}{B}
    \caption{Structural decomposition of a typical avoider of 231}
    \label{fig:decompmax}
\end{figure}

The elements to the left of the maximum, \(A\), have the structure of a \(231\)
avoiding permutation, and the elements to the right of the maximum, \(B\), have the
structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
\emph{lower-left region} and \(B\) the \emph{upper-right region}.

@@

\section{Wilf-classes with patterns of length 1.}
When considering the mesh patterns of length \(2\) it will be useful to know
the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
\(\av{\perm{2,3,1}}\), this means that we are sonsdering the set \(\av{\{231,p\}}\)
where \(p\) is a mesh-pattern.

It can be seen that the patterns in the following set are set equivalent.
\begin{equation*}
    \left\{
        \begin{array}{c}
        \textpattern{}{1}{},
        \textpattern{}{1}{0/0},
        \textpattern{}{1}{1/0},
        \textpattern{}{1}{0/0,1/0},
        \textpattern{}{1}{0/1},\\
        \textpattern{}{1}{0/0,0/1},
        \textpattern{}{1}{1/1},
        \textpattern{}{1}{1/0,1/1},
        \textpattern{}{1}{0/1,1/1}
    \end{array}
    \right\}
\end{equation*}
This is due to the fact that every permutation except the empty
permutation must contain an occurrence of all of these patterns.

@@

\nextvar
The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
quadrant marked mesh patterns studied by Remmel,
Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
avoiders by decomposing a typical avoider of \(p_1\) around the maximum element
in order to give the following structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\). The lower left region however can create occurrences
of \(p\) and therefore must also avoid \(p\) as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

@@

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of an avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with it's maximum, by not allowing the left part of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.

@@

Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
pattern are permutations that do not start with their minimum. In this case we
perform the decomposition around the leftmost element
\begin{equation*}
    \mathcal{\var}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
All of these classes have the same generating function namely
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1)
\end{equation*}
Enumeration of this generating function gives
\begin{equation*}
    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
\end{equation*}

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(r\) must fit into the following form precisely once.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function can be enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\end{equation*}

@@

In order to consider symmetries we must only take the symmetries that preserve
the pattern \(\perm{2,3,1}\). If we take any of the symmetries alone the
permutation is different. The only symmetry that preserves the pattern
\(\perm{2,3,1}\) is that of \emph{reverse-complement-inverse}. Using this
symmetry to merge classes gives us \(61\) classes of trivial
Wilf-equivalences.

Computing avoiders up to length \(10\) gives us \(23\) Wilf-classes,
of which \(13\) are non-trivial.

@@

\begin{note}
    \label{not:setequiv}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\)
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). \cite{DBLP:journals/ejc/Claesson01}
\end{note}
\begin{proof}
    Let \(\pi\) be a permutation in \(\av[n]{q}\) take the permutation
    in one-line notation and insert a dash between each ascent in \(\pi\). This
    corresponds to set partitions where the blocks are the elements between the
    dashes, the blocks are listed in increasing order of their least element,
    with the elements written in each block in descending order.
\end{proof}

\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{1,2,4,5\},\{3\},\{6,9\},\{8,7\}\}\).
\end{example}

We are looking at permutations in \(\av{\perm{2,3,1}}\), all of these permutations
also avoid the mesh pattern in \NoteRef{not:setequiv}.

@@

\subsection{}
The containers of the following patterns
have exactly one occurrence.
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Once again we consider the regions corresponding to the unshaded regions, For each
pattern
except the first the two regions are independent, and one contains
any avoider of \(\perm{2,3,1}\) and the other must contain a
decreasing sequence by \LemmaRef{lem:incdec}.
Let us consider the first pattern separately. In order to
avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
we can add some additional restrictions
\begin{equation}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
        \draw (1.5,0) -- (1.5,1)
              (1.5,2) -- (1.5,3);
        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \end{tikzpicture}
    }
    \label{eqn:splitboxes}
\end{equation}
Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
two do not interact in any manner. The containers of this pattern are counted
the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
The containers
have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
therefore gives us
\begin{equation*}
    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
\end{equation*}

\subsection{}
The set/symmetry classes containing the following patterns are
Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not set equivalent or
symmetries of each other
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/2,
                    1/0,1/1,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/0,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
In this case it is better to consider the containers of the patterns
instead of the avoiders due to the amount of shadings in the mesh.

We look at the containers of the pattern \(m_1\), there can only ever
be one occurrence of this pattern in a permutation corresponding to
the last point in the permutation and the minimum.Consider an occurrence of \(m_1\),
the points in the two regions corresponding to the the two boxes must form decreasing
subsequences.
For a permutation of length \(k\) if we fix the number of points in
one of the boxes the number of points in the other box is determined.
Therefore we can have any number of points from \(\setrange{0}{k-2}\)
points in the bottom box. Therefore there are \(k-1\) containers of
length \(k\). These permutations correspond to set partitions of \(k\)
points into exactly two non-overlapping parts partitioned by the first
element and the minimum.

@@

Therefore these two patterns have been shown to have the same number of
avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
The avoiders have enumeration
\begin{equation*}
    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc
\end{equation*}

@@

First we consider the coincidence in two parts then consolidate these parts.
Consider the coincidence shown in \eqref{eq:singleud}. The easiest way to
show that these are equinumerous is to consider the containers in the realm
of set partitions.

Considering an occurrence of either of these patterns in a permutation by
the mesh sets we know the following about the points corresponding
to the points in the patterns.
\begin{itemize}
    \item The point corresponding to the first point in both patterns
        must lie in the first block of our set partition (there are
        no points southwest from it in the permutation).
    \item The point corresponding to the second point in both patterns
        is a block bottom (there are no points southeast of it in
        the permutation).
    \item If the region corresponding to box \((2,2)\) in an occurrence of
    \(m_1\) is empty, then the point corresponding to the second point is
    precisely the last block bottom. If the region corresponding to box \((0,1)\)
    in an occurrence of \(m_2\) is empty, then the point corresponding to the
    second point is precisely the first block bottom.
    If these regions are non-empty then the block containing the point
    corresponding to the second point in both patterns contains only the point
    (it is a singleton block).
\end{itemize}
This tells us that an occurence of the patterns must happen when there
is a singleton block occurring after the first block. The difference
between the patterns is in the underlying classical pattern. This
means that permtuations containing \(m_1\) correspond to set partitions
with a singleton block with value one higher than some element in the
block containing \(1\). The permutations containing \(m_2\) correspond
to the set partitions containing a block with block bottom having value
one lower than some element in the block containing \(1\) and if this
block is not the block containing \(1\) then it is a singleton block.

@@

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\perm{2,3,1},m_1}\) we gain the following.
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore all four patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc
\end{equation*}
This corresponds to OEIS sequence number \href{https://oeis.org/A035929}{A035929}
with offset 1.

@@

First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\)
we can perform the usual structural decomposition of an avoider
of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
% Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
% precisely once.
If \(\scriptvar_1\) set \(\av{\perm{2,3,1},m_1}\) then any permutation in
\(\scriptvar_1\) either starts with a maximum or does not, giving us the
decomposition
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}

@@

First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
around the maximum, we have different conditions if we start with the maximum or not.
\begin{equation*}
    \scriptvar_1 =
    \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
\end{equation*}

>>>>

Another question often asked in the field of permutation patterns is that
of Wilf-equivalence. Two patterns \(\pi\) and \(\sigma\) are said to be
Wilf-equivalent if their avoidance sets have the same size at each
length. More formally:

\begin{definition}[Wilf-equivalence]
    Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}
    if for all \(k_{} \ge 0\), \(\size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\).
    Two sets of permutation patterns \(R\) and \(S\) are are
    \emph{Wilf-equivalent} if for all \(k_{} \ge 0\),
    \(\size{\av[k]{R}} = \size{\av[k]{S}}\).
\end{definition}

Wilf-equivalence is of interest since if two permutation classes are enumerated
in the same way then there should exist a bijection between them, and therefore
any other combinatorial object that they represent.

@@

There are a number of symmetries we can use when examining Wilf-equivalences
to reduce the amount of work, it can be easily seen that the reverse, complement
and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and therefore these classes are trivially
Wilf-equivalent.
\begin{figure}[!htb]
\begin{align*}
    \rev{\textpattern{}{2,3,1}{}} &= \textpattern{}{1,3,2}{}\\
    \com{\textpattern{}{2,3,1}{}} &= \textpattern{}{2,1,3}{}\\
    \inv{\textpattern{}{2,3,1}{}} &= \textpattern{}{3,1,2}{}
\end{align*}
\caption{The operations reverse, complement and inverse for the pattern 231}
\label{fig:symm}
\end{figure}

The group of symmetries on permutations is isomorphic to the dihedral group
of order \(8\), the group of symmetries of a square. \emph{Reverse-inverse} and
\emph{reverse} correspond can be taken as generators of the dihedral group.

Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
must only use these symmetries when they preserve the dominating pattern, if we
were to allow other symmetries, then the equivalences calculated in the previous
section do not necessarily hold.

Throughout this section we will consider Wilf-equivalences of patterns
whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\). We will
use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
be the usual Catalan generating function satisfying \(C(x) = 1 + C(x)^2\).
This is easy to see by structural decomposition around the maximum, as
shown in \FigureRef{fig:decompmax}.

\begin{figure}[!ht]
    \centering
    \decompmax{0/1,1/0}{A}{}{}{B}
    \caption{Structural decomposition of a non-empty avoider of 231}
    \label{fig:decompmax}
\end{figure}

The elements to the left of the maximum, \(A\), have the structure of a \(231\)
avoiding permutation, and the elements to the right of the maximum, \(B\), have the
structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
\emph{lower-left section} and \(B\) the \emph{upper-right section}.

@@

\section{Wilf-classes with patterns of length 1.}
When considering the mesh patterns of length \(2\) it will be useful to know
the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
\(\av{\perm{2,3,1}}\), this means that we are consdering the set \(\av{\{231,p\}}\)
where \(p\) is a mesh-pattern of length 1.

The patterns in the following set are set equivalent,
\begin{equation*}
    \left\{
        \begin{array}{c}
        \textpattern{}{1}{},
        \textpattern{}{1}{0/0},
        \textpattern{}{1}{1/0},
        \textpattern{}{1}{0/0,1/0},
        \textpattern{}{1}{0/1},\\
        \textpattern{}{1}{0/0,0/1},
        \textpattern{}{1}{1/1},
        \textpattern{}{1}{1/0,1/1},
        \textpattern{}{1}{0/1,1/1}
    \end{array}
    \right\}
\end{equation*}
due to the fact that every permutation, except the empty
permutation must contain an occurrence of all of these patterns.

@@

\nextvar
The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
quadrant marked mesh patterns studied by Remmel,
Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
avoiders by decomposing a non-empty avoider of \(p_1\) around the maximum element
in order to give the following structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper-right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\). The lower-left section however can create occurrences
of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

@@

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of a non-empty avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum:
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with it's maximum, by not allowing the lower-left section of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.

@@

Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
pattern are permutations that do not start with their minimum. In this case we
perform the decomposition around the leftmost element
\begin{equation*}
    \mathcal{\var}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
All of these classes have the same generating function, namely
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1).
\end{equation*}
Enumeration of this generating function gives
\begin{equation*}
    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset}
\end{equation*}

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(r\) must fit into precisely on of the following twp forms.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, minimum cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function is enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364} }
\end{equation*}

@@

In order to consider symmetries we must only take the symmetries that preserve
the pattern \(\perm{2,3,1}\). The only symmetry that preserves the pattern
\(\perm{2,3,1}\) is \emph{reverse-complement-inverse}. Using this
symmetry to merge classes gives us \(61\) classes of trivial
Wilf-equivalences, these Wilf-equivalences are explained by patterns being either
coincident, or being the reverse-complement-inverse of a pattern.

Computing avoiders up to length \(10\) suggest that there are at least \(23\)
Wilf-classes, of which \(13\) are non-trivial.

@@

\begin{note}
    \label{not:setequiv}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01})
\end{note}
\begin{proof}
    Let \(\pi\) be a permutation in \(\av[n]{q}\)
    in one-line notation and insert a dash between each ascent in \(\pi\). This
    corresponds to set partitions where the blocks are the elements between the
    dashes, the blocks are listed in increasing order of their least element,
    with the elements written in each block in descending order.
\end{proof}

\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
\end{example}

We call the least element in each block the \emph{block bottom}

We are looking at permutations in \(\av{\perm{2,3,1}}\), all of these permutations
also avoid the mesh pattern in \NoteRef{not:setequiv}, \ie
\(\av{\perm{2,3,1}} \subset \av{q}\).

@@

\subsection{}
The containers of the following patterns
have exactly one occurrence.
\begin{equation*}
    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
\end{equation*}
Once again we consider the regions corresponding to the unshaded regions, For each
pattern
except the first the two regions are independent, and one contains
any avoider of \(\perm{2,3,1}\) and the other must contain a
decreasing sequence by \LemmaRef{lem:incdec}.
Let us consider the first pattern separately. In order to
avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
we can add some additional restrictions, \ie all elements in the top region must
be to the right of all elements in the bottom region.
\begin{equation}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
        \draw (1.5,0) -- (1.5,1)
              (1.5,2) -- (1.5,3);
        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
    \end{tikzpicture}
    }
    \label{eqn:splitboxes}
\end{equation}
Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
two do not interact in any manner. The containers of this pattern are counted
the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
The containers
have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
therefore gives us
\begin{equation*}
    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
\end{equation*}

\subsection{}
\nextvar
The trivial Wilf-classes containing the following patterns are
Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not coincident or
symmetries of each other
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0,0/2,
                    1/0,1/1,
                    2/0,2/1,2/2}
    \text{ and }
    m_2 = \pattern{}{1,2}{0/0,
                    1/0,1/1,1/2,
                    2/0,2/1,2/2}
\end{equation*}
In this case it is better to consider the containers of the patterns
instead of the avoiders due to the amount of shadings in the mesh.

We look at the containers of the pattern \(m_1\), there can only ever
be one occurrence of this pattern in a permutation corresponding to
the last point in the permutation and the minimum. Consider an occurrence of \(m_1\),
the points in the two regions corresponding to the the two boxes must form decreasing
subsequences.
For a permutation of length \(k\) if we fix the number of points in
one of the boxes the number of points in the other box is determined.
Therefore we can have any number of points from \(\setrange{0}{k-2}\)
points in the bottom box. Therefore there are \(k-1\) containers of
length \(k\). These permutations correspond to set partitions of \(k\)
points into exactly two non-overlapping parts partitioned by the first
element and the minimum.

@@

Therefore these two patterns have been shown to have the same number of
avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
The avoiders have general form
\begin{equation*}
  \var_k = C_k - (k-1), \var_0 = 1
\end{equation*}
and have enumeration
\begin{equation*}
    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc
\end{equation*}

@@

First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
\eqref{eq:singleud}. The easiest way to
show that these are equinumerous is to consider the containers as set partitions.

Considering an occurrence of either of these patterns in a permutation  we know
the following about the points corresponding to the points in the patterns.
\begin{itemize}
    \item The point corresponding to the first point in both patterns
        must lie in the first block of the set partition (there are
        no points southwest from it in the permutation).
    \item The point corresponding to the second point in both patterns
        is a block bottom (there are no points southeast of it in
        the permutation).
    \item If the region corresponding to box \((2,2)\) in an occurrence of
    \(m_1\) is empty, then the point corresponding to the second point is
    precisely the last block bottom. If the region corresponding to box \((0,1)\)
    in an occurrence of \(m_2\) is empty, then the point corresponding to the
    second point is precisely the first block bottom.
    If these regions are non-empty then the block containing the point
    corresponding to the second point in both patterns contains only the point
    (it is a singleton block).
\end{itemize}
This tells us that an occurence of the patterns must happen when there
is a singleton block occurring after the first block. The difference
between the patterns is in the underlying classical pattern. This
means that permutations containing \(m_1\) correspond to set partitions
with a singleton block with value one higher than some element in the
block containing \(1\). The permutations containing \(m_2\) correspond
to the set partitions containing a block with block bottom having value
one lower than some element in the block containing \(1\) and if this
block is not the block containing \(1\) then it is a singleton block. This
proves that the containers of both of these patterns in \(\av{231}\) are
equinumerous, and therefore so are their avoiders.

@@

Now we must consolidate these two subclasses. In order to do this we must
consider the decomposition around the leftmost point of a permutation in
\(\av{\perm{2,3,1},m_1}\) we gain the following.
\begin{equation*}
    \scriptvar_3 = \varepsilon \sqcup
    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
\end{equation*}
It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
have the same enumeration, and therefore all four patterns are Wilf-equivalent
in \(\av{\perm{2,3,1}}\) with generating function satisfying
\begin{equation*}
    \var(x) = 1 + x\var(x)\varmaxl(x)
\end{equation*}
This can be enumerated to give the sequence
\begin{equation*}
    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc \tag{\href{https://oeis.org/A035929}{OEIS: A035929} offset 1}
\end{equation*}

@@

First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\).
We can perform the usual structural decomposition of an avoider
of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
% Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
% precisely once.
If \(\scriptvar_1\) is the set \(\av{\perm{2,3,1},m_1}\) then any permutation in
\(\scriptvar_1\) either starts with a maximum or does not, giving us the
decomposition
\begin{equation*}
    \scriptvar_1 = \varepsilon \sqcup
    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
\end{equation*}

@@

First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
around the maximum. We have different conditions if we start with the maximum or not.
\begin{equation*}
    \scriptvar_1 =
    \varepsilon \sqcup
    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
    \sqcup
    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
\end{equation*}

Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
conditions for starting with the maximum
\begin{equation*}
    \scriptvar_2 =
    \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
    \sqcup
    \decompleft{0/1,1/0,1/1}{\scriptvar_2}{}{}{}
\end{equation*}
This gives us two generating functions satisfying the following pair of equations
\begin{align}
    \var_1(x) &= 1 + xC(x)(\varmaxl(x)-1) +x\var_1(x) \label{eqn:genL1}\\
    \text{ and }\var_2(x) &= 1 + x(C(x)-1)^2 + x\var_2(x) \label{eqn:genL2}
\end{align}
In order for these two functions to give the same value it is necessary to show that
\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
\(C(x)(\varmaxl(x)-1) = (C(x)-1)^2\).
\begin{align*}
    &&(C(x)-1)^2 &= C(x)(\varmaxl(x)-1)\\
        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(\varmaxl\) and \(C\)}\\
        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
\end{align*}
The final line is always satisfied since it is the form of \(C(x)\), and
therefore the two generating functions are equal.
